Estimating R Anxiety Level Distribution Among Students At the University Using MRP

### 2020-12-16 outputs/paper/paper.Rmd
Add files via upload
Github Desktop not working, committing paper.rmd and paper.pdf manually.

---
title: "Estimating R Anxiety Level Distribution Among Students At the University Using MRP"
author: "Yiqu Ding"
thanks: "Code and data are available at: https://github.com/dding33/STA304-PS5."
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
abstract: |
  In this report, I look at different factors that affect a studebt's anxiety level towards   using R. The data is obtained from onlien survey responses which contained variables that   could influence anxiety levels. I then run a multilevel regression on the sample and       post-stratify them using a simulated student census. After the post-stratification, we 
  see the estimated mean anxiety scores are higher average anxiety scores from the raw 
  data.

@@

  Relatively speaking, the science of statistics is a new discipline. In 1998, the public image of statistics was poor, and almost nobody knows what statisticians do[from stats]. Now, statistics is an essential tool for nearly all millennial industries. Accompanied by technological improvements in computers, R has become a necessary tool for all statistical practitioners; that makes the teaching and training of R extremely important.

  With that being said, R’s mastering has not been on student's to-do list until recent years. At the University of Toronto, up until fall in 2018, R’s learning is not compulsory until third-year courses. Many students expressed surprise when they first see R’s use in the classroom and are confused. The anxiety issue persists three years after the department made STA130 compulsory, which was an introduction to statistics and R. Studies show that the anxiety level affects students’ performance in the classroom and has the potential for further investigation[onlineanxiety].

@@

  In paper we run our analysis in `R` [citeR]. We also use the `tidyverse` package which was written by tydr.

@@

The total sample size is 48, from which 7 respondents answered ‘no’ to the qualification questions. That makes the sample size 41. Figure x shows the distribution of the programs from the sample. The coloring at the end of the bars indicates the respondent does not have any coding experience. We notice that this is a small part of the sample.

Figure x displays the anxiety score distribution from the raw data set. We see two prominent peaks in the distribution: around 3, which indicates the respondents do not feel very anxious, and around 7.5 indicates the respondents feel quite anxious. Most responses fall between these two peaks, with few respondents(3 out of 41) reports extreme anxiety scores towards 1 or 10. Looking at Figure x, we notice that our sample does not contain any students with a cumulative GPA lower than C. This skewness means that our sample is biased; specifically, students with higher cgpa have a stronger incentive to answer the survey.  We will adjust for this in the model by incorporating random effects.

We must point out that studies show the response biases for sensitive topics center are near zero, but the responses are unreliable or noisy[sensitive_info]. Since the cgpa and the anxiety score reveal information about students’ academic behavior, we consider them sensitive topics. We follow steps from [mathexample] using the `brms` package[brms1, brms2] to adjust for this. Further explanation will continue in the Model section.

```{r}
#Import the data and attach the variables to the path
clean_data <- read_csv("~/GitHub/STA304-PS5/inputs/data/clean_data.csv")
attach(clean_data)
```

@@

We use ladder four to create post-stratification cells. They contain information about an individual to divide them into groups and identify each individual using these four variables. The four variables make 2*8*5*2 = 160 possible cells. The distribution of each variable is simulated based on a rough estimation of the population. See fig x and x for a summary of the census. This report’s results are not estimates of the University of Toronto’s actual R anxiety distribution, even though the sample is collected from real respondents.

@@

```{r, fig.cap = "Program Distribution in Census"}
census <- read_csv("~/GitHub/STA304-PS5/inputs/data/census.csv")
#Create graph for program distribution in the census
ggplot(census, aes(x = program)) +
  geom_bar(alpha = 0.8, color = "cyan4", fill = "cyan4") +
  theme_minimal() +
  xlab("Programs")+ylab("Number of Respondents")+
  coord_flip()
```

@@

\begin{equation}
\hat{y}_S^{PS} = \frac{\sum{N_j\hat{y_j}}}{\sum{N_j}}  (\#eq:PS)
\end{equation}

@@

\begin{equation}
\hat{y} = \beta_0 + \beta_{sex}d_{sex} + \beta_{cgpa}x_{cgpa} + \beta_{pro}x_{pro} + \beta_{code}x_{code} + e (\#eq:formula)
\end{equation}

@@

```{r, echo = FALSE}
#Perform 3-fold cross validation
kfold <- kfold(mod1, K = 3, save_fits = TRUE, seed=TRUE)
#Define a error calculation function
rmse <- function(y, yrep) {
  yrep_mean <- colMeans(yrep)
  sqrt(mean(yrep_mean - y)^2)
}
#Predict ressponses and estimate the error
kfp <- kfold_predict(kfold)
rmse(y = kfp$y, yrep = kfp$yrep)
```

@@

  We use `add_predicted_draws()` to come up with estimations and their 95% confidence intervals. Figures 4-7 show the prediction results by different sex, program, cgpa, and coding experience, with the raw data results. We see that the MRP estimates produce a higher mean anxiety score comparing to the raw results. Specifically,

@@

```{r, fig.cap="MRP estimates vs Raw data in different sex groups"}
sex_prop <- read_csv("~/GitHub/STA304-PS5/inputs/data/sex_prop.csv")
res_sex <- mod1 %>% 
  add_predicted_draws(newdata = sex_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% #caculate the prediction
  group_by(sex, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% #Aggregrate the results
  group_by(sex) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)#Produce the 95% confidence interval 
  )

@@

```{r, fig.cap="MRP estimates vs Raw data in different programs"}
program_prop <- read_csv("~/GitHub/STA304-PS5/inputs/data/program_prop.csv")
res_program <- mod1 %>% 
  add_predicted_draws(newdata = program_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% 
  group_by(program, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% 
  group_by(program) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)
  )

@@

```{r, fig.cap="MRP estimates vs Raw data within different cgpa groups"}
cgpa_prop <- read_csv("~/GitHub/STA304-PS5/inputs/data/cgpa_prop.csv")
res_cgpa <- mod1 %>% 
  add_predicted_draws(newdata = cgpa_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% 
  group_by(cgpa, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% 
  group_by(cgpa) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)
  )

@@

```{r, fig.cap="MRP estimates vs Raw data with different coding experiences"}
code_prop <- read_csv("~/GitHub/STA304-PS5/inputs/data/code_prop.csv")
res_code <- mod1 %>% 
  add_predicted_draws(newdata = code_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% 
  group_by(coding_exp, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% 
  group_by(coding_exp) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)
  )

@@

[intro] states that MRP works well adjusting for biased samples only if the under/over-represented variables are present in the post-stratification data set. With that being said, to get a more reliable prediction result, future analysis can contain a pre-analysis which has more variables and use stepwise regression to select variables that contribute to the accuracy of the model.

>>>>

---
title: "Estimating R Anxiety Level Distribution Among Students At the University Using MRP"
author: "Yiqu Ding"
thanks: "Code and data are available at: https://github.com/dding33/STA304-PS5."
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
abstract: |
  In this report, I look at different factors that affect a student's anxiety level towards using R. The data is obtained from online survey responses, which contained variables that could influence anxiety levels. I then run a multilevel regression on the sample and post-stratify them using a simulated student census. After the post-stratification, we see the estimated mean anxiety scores are higher than average anxiety scores from the raw data.

@@

  Relatively speaking, the science of statistics is a new discipline. In 1998, the public image of statistics was poor, and almost nobody knows what statisticians do[@fromstats]. Now, statistics is an essential tool for nearly all millennial industries. Accompanied by technological improvements in computers, R has become a necessary tool for all statistical practitioners; that makes the teaching and training of R extremely important.

  With that being said, R’s mastering has not been on student's to-do list until recent years. At the University of Toronto, up until fall in 2018, R’s learning is not compulsory until third-year courses. Many students expressed surprise when they first see R’s use in the classroom and are confused. The anxiety issue persists three years after the department made STA130 compulsory, which was an introduction to statistics and R. Studies show that the anxiety level affects students’ performance in the classroom and has the potential for further investigation[@onlineanxiety].

@@

  In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` package which was written by [@tydr].

@@

The total sample size is 48, from which 7 respondents answered ‘no’ to the qualification questions. That makes the sample size 41. Figure 1 shows the distribution of the programs from the sample. The coloring at the end of the bars indicates the respondent does not have any coding experience. We notice that this is a small part of the sample.

Figure 3 displays the anxiety score distribution from the raw data set. We see two prominent peaks in the distribution: around 3, which indicates the respondents do not feel very anxious, and around 7.5 indicates the respondents feel quite anxious. Most responses fall between these two peaks, with few respondents(3 out of 41) reports extreme anxiety scores towards 1 or 10. Looking at Figure 2, we notice that our sample does not contain any students with a cumulative GPA lower than C. This skewness means that our sample is biased; specifically, students with higher cgpa have a stronger incentive to answer the survey.  We will adjust for this in the model by incorporating random effects.

We must point out that studies show the response biases for sensitive topics center are near zero, but the responses are unreliable or noisy[@sensitiveinfo]. Since the cgpa and the anxiety score reveal information about students’ academic behavior, we consider them sensitive topics. We follow steps from [@mathexample] using the `brms` package[@brmone, @brmtwo] to adjust for this. Further explanation will continue in the Model section.

```{r}
#Import the data and attach the variables to the path
clean_data <- read_csv("~/Downloads/STA304-PS5-master/inputs/data/clean_data.csv")
attach(clean_data)
```

@@

We use ladder four to create post-stratification cells. They contain information about an individual to divide them into groups and identify each individual using these four variables. The four variables make 2*8*5*2 = 160 possible cells. The distribution of each variable is simulated based on a rough estimation of the population. See fig 4 and 5 for a summary of the census. This report’s results are not estimates of the University of Toronto’s actual R anxiety distribution, even though the sample is collected from real respondents.

@@

```{r, fig.cap = "Program Distribution in Census"}
census <- read_csv("~/Downloads/STA304-PS5-master/inputs/data/census.csv")
#Create graph for program distribution in the census
ggplot(census, aes(x = program)) +
  geom_bar(alpha = 0.8, color = "cyan4", fill = "cyan4") +
  theme_minimal() +
  xlab("Programs")+ylab("Number of Respondents")+
  coord_flip()
```

@@

\begin{equation}
\hat{y}_S^{PS} = \frac{\sum{N_j\hat{y_j}}}{\sum{N_j}}  
\end{equation}

@@

\begin{equation}
\hat{y} = \beta_0 + \beta_{sex}d_{sex} + \beta_{cgpa}x_{cgpa} + \beta_{pro}x_{pro} + \beta_{code}x_{code} + e 
\end{equation}

@@

```{r, echo = FALSE, include = FALSE}
#Perform 3-fold cross validation
kfold <- kfold(mod1, K = 3, save_fits = TRUE, seed=TRUE)
#Define a error calculation function
rmse <- function(y, yrep) {
  yrep_mean <- colMeans(yrep)
  sqrt(mean(yrep_mean - y)^2)
}
#Predict ressponses and estimate the error
kfp <- kfold_predict(kfold)
rmse(y = kfp$y, yrep = kfp$yrep)
```

@@

  We use `add_predicted_draws()` to come up with estimations and their 95% confidence intervals. Figures 6-9 show the prediction results by different sex, program, cgpa, and coding experience, with the raw data results. We see that the MRP estimates produce a higher mean anxiety score comparing to the raw results. Specifically,

@@

```{r, fig.cap="MRP estimates vs Raw data in different sex groups"}
sex_prop <- read_csv("~/Downloads/STA304-PS5-master/inputs/data/sex_prop.csv")
res_sex <- mod1 %>% 
  add_predicted_draws(newdata = sex_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% #caculate the prediction
  group_by(sex, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% #Aggregrate the results
  group_by(sex) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)#Produce the 95% confidence interval 
  )

@@

```{r, fig.cap="MRP estimates vs Raw data in different programs"}
program_prop <- read_csv("~/Downloads/STA304-PS5-master/inputs/data/program_prop.csv")
res_program <- mod1 %>% 
  add_predicted_draws(newdata = program_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% 
  group_by(program, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% 
  group_by(program) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)
  )

@@

```{r, fig.cap="MRP estimates vs Raw data within different cgpa groups"}
cgpa_prop <- read_csv("~/Downloads/STA304-PS5-master/inputs/data/cgpa_prop.csv")
res_cgpa <- mod1 %>% 
  add_predicted_draws(newdata = cgpa_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% 
  group_by(cgpa, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% 
  group_by(cgpa) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)
  )

@@

```{r, fig.cap="MRP estimates vs Raw data with different coding experiences"}
code_prop <- read_csv("~/Downloads/STA304-PS5-master/inputs/data/code_prop.csv")
res_code <- mod1 %>% 
  add_predicted_draws(newdata = code_prop, allow_new_levels = TRUE) %>% 
  rename(anxiety_predict = .prediction) %>% 
  mutate(anxiety_predict_prop = anxiety_predict*prop) %>% 
  group_by(coding_exp, .draw) %>% 
  summarise(anxiety_predict = sum(anxiety_predict_prop)) %>% 
  group_by(coding_exp) %>% 
  summarise(
    mean = mean(anxiety_predict),
    lower = quantile(anxiety_predict, 0.025),
    upper = quantile(anxiety_predict, 0.975)
  )

@@

[@intro] states that MRP works well adjusting for biased samples only if the under/over-represented variables are present in the post-stratification data set. With that being said, to get a more reliable prediction result, future analysis can contain a pre-analysis which has more variables and use stepwise regression to select variables that contribute to the accuracy of the model.
